---
output:
  rmdformats::html_docco:
    fig_width: 6
    fig_height: 6
    thumbnails: false
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## CHC confirmatory factor analysis {.tabset}

### Data preparation 

The following packages were used: 



```{r, results = 'hide', echo = T, message = F}
library(haven)
library(tidyverse)
library(lavaan)
library(psych)
library(semTools)
library(mice)

```

```{r, include=F}

arcli <- read_sav("C:\\Users\\lachy\\Dropbox\\ARCLI data entry\\BLINDED DATA SETS\\BLINDED _DO NOT TOUCH\\ARCLI_V1_Baseline_ALL_13-JUL-2021.sav")
``` 
 

```{r, include = F}

arcli <- na_if(arcli, -999) # mark -99's as missing data

arcli <- na_if(arcli, -99)

# Rename problematic variable names

arcli <- arcli %>% 
  rename(CDR_ImmWRCorrect_v1 = `CDR_ImmWRCorrect#_v1`)

arcli <- arcli %>% 
  rename(CDR_DelayedWRCorrect_v1 = `CDR_DelayedWRCorrect#_v1`)


# create new missing overall score

arcli$CDR_NumericWMOverallAcc_v1 <- (arcli$CDR_NumericWMOrigStimAcc_v1 + arcli$CDR_NumericWMNewStimAcc_v1) / 2

arcli$Sex <- as.factor(arcli$Sex)

levels(arcli$Sex) <- c("male", "female")
```


Select the relevant cognitive test variables from the data and remove blank (empty) rows:


```{r}

cog_vars <- arcli %>% 
  select(Sex, IDno, CDR_RVIPRTms_v1, Initialage,CDR_RVIPacc_v1,TMT_Bms_v1, CDR_ImmWRCorrect_v1,
         CDR_DelayedWRCorrect_v1, CDR_SimpleRTms_v1, CDR_ChoiceRTms_v1,       
         CDR_NumericWMOverallAcc_v1, CDR_DigitVigRTms_v1, SCB_ConStroopRTms_v1,        
         SCB_InconStroopRTms_v1, Jenson8ChoiceDecisionTimems_v1,
         Jenson4ChoiceDecisionTimems_v1,           
         Jenson1ChoiceDecisionTimems_v1,WASImatrixreasoningrawscore, 
         WASI_vocabrawscore, MMSE, Serial3sAcc_v1, Serial7sAcc_v1, TMT_Ams_v1)


# Remove rows without cog results

cog_vars <- cog_vars[which(rowMeans(is.na(cog_vars)) < 0.80),]

```


The reaction time variables are log2 transformed and plotted in histograms to assess for errors/outliers:
 

```{r}

RT_vars <- cog_vars %>% select(
  CDR_RVIPRTms_v1,TMT_Bms_v1,CDR_SimpleRTms_v1, CDR_ChoiceRTms_v1, 
  CDR_DigitVigRTms_v1, SCB_ConStroopRTms_v1, SCB_InconStroopRTms_v1, 
  Jenson8ChoiceDecisionTimems_v1,Jenson4ChoiceDecisionTimems_v1,
  Jenson1ChoiceDecisionTimems_v1, TMT_Ams_v1) %>% 
  names() # reaction time measures 

cog_vars <- cog_vars %>% 
  mutate(across(all_of(RT_vars), ~ log2(.), .names = "log{.col}")) # log 2 transform


# display histograms 

cog_vars[,c(24:34)] %>% select(where(is.numeric)) %>% 
  multi.hist(global = F)

cog_vars %>% select(CDR_ImmWRCorrect_v1, CDR_DelayedWRCorrect_v1,
                               CDR_NumericWMOverallAcc_v1, MMSE,
                               WASImatrixreasoningrawscore, WASI_vocabrawscore,
                               Serial3sAcc_v1, Serial7sAcc_v1) %>% 
  multi.hist(global = F)


```


Outliers are then winsorised at 3 standard deviations to reduce the excessive influence of problematic and unlikely values (e.g, indicating a misunderstood test): Histograms of the winsorised variables are then plotted. 


```{r}

# winsorise function

winsorise <- function(x, sd){
  maxin <- max(x[scale(x) < sd],na.rm=T)
  minin <- min(x[scale(x) > -sd],na.rm=T)
  x <- ifelse(x > maxin & !is.na(x), maxin, x)
  x <- ifelse(x < minin & !is.na(x), minin, x)
  x
}

# Winsorise outliers 

cog_vars <- cog_vars %>% # winsorise everything numeric except age
  mutate(across(c(where(is.numeric), -Initialage), winsorise, sd = 3, .names = "w{.col}"))

cog_vars2 <- cog_vars %>% select(IDno, Initialage, Sex, starts_with("wlog"), 
                                 wWASImatrixreasoningrawscore, wWASI_vocabrawscore, 
                                 wCDR_ImmWRCorrect_v1,wCDR_DelayedWRCorrect_v1, 
                                 wCDR_NumericWMOverallAcc_v1, wMMSE,wSerial3sAcc_v1,
                                 wSerial7sAcc_v1) # select relevant variables

# histograms of winsorised variables

cog_vars2 %>% select(starts_with("wlog")) %>% 
  multi.hist(global = F)

cog_vars2 %>% select(wCDR_ImmWRCorrect_v1, wCDR_DelayedWRCorrect_v1,
                    wCDR_NumericWMOverallAcc_v1, wMMSE,
                    wWASImatrixreasoningrawscore, wWASI_vocabrawscore,
                    wSerial3sAcc_v1, wSerial7sAcc_v1) %>% 
  multi.hist(global = F)


```


Rows with more than 50% missing data for the cognitive test are dropped and the percentage of missing data per variable is calculated: 

```{r}

# delete rows with > 50% missing data 

cog_vars2 <- cog_vars2[which(rowMeans(is.na(cog_vars2)) < 0.50),]

n_miss <- colSums(is.na(cog_vars2)) # check missing data per variable

sort(n_miss / nrow(cog_vars2)) 

median(n_miss / nrow(cog_vars2))
```

Reverse score reaction time measures and rescale variables to more interpretable units:


```{r}

reverse_RT <- function(x){
  x <- x * -1
  x <- x + max(abs(x),na.rm=T) + 1
  x
}

cog_vars2 <- cog_vars2 %>% 
  mutate(across(starts_with("wlog"), reverse_RT, .names = "{.col}_rev"))


## rescale accuracy variables (10% units)

cog_vars2 <- cog_vars2 %>% mutate(across(starts_with("wSerial"), ~  . / 10))

cog_vars2 <- cog_vars2 %>% mutate(wCDR_NumericWMOverallAcc_v1 = wCDR_NumericWMOverallAcc_v1 / 10)

# rescale WASI scores to per 10 points 

cog_vars2 <- cog_vars2 %>% 
  mutate_at(vars("wWASImatrixreasoningrawscore",
                 "wWASI_vocabrawscore"),
            .funs = ~ . / 10)
```


### imputation

Perform single imputation to replace missing values in cognitive test items using predictive mean matching. Only variables with a correlation greater than 0.10 are used to impute missing data for each variable. Age and gender are included in each imputation model. 


```{r, message = F, quietly = T}

cog_vars3 <- cog_vars2 %>% 
  select(IDno, Sex, Initialage, ends_with("_rev"), wCDR_ImmWRCorrect_v1, 
         wWASI_vocabrawscore, wWASImatrixreasoningrawscore,
         wCDR_DelayedWRCorrect_v1, wCDR_NumericWMOverallAcc_v1, wMMSE,
         wSerial3sAcc_v1, wSerial7sAcc_v1) # select variables for imputation

## single imputation

# include age + gender and any cog var with > |0.10| correlation in 
# imputation model 

predictormatrix <- quickpred(cog_vars3, 
                           include=c("Sex", "Initialage"),
                           exclude="IDno",
                           mincor = 0.1) 

cog_vars3_imp <- mice(cog_vars3, m = 1, method = "pmm",
                      predictorMatrix = predictormatrix, diagnostics = T, print =F,
                      seed = 10) # single imputation

imp_dat <- complete(cog_vars3_imp) # imputed data 

```


Check that the imputation produced reasonable replacements. The red circles are the imputed observations. Imputation 1 is the original data, imputation 2 is the data with imputations. 

```{r}

stripplot(cog_vars3_imp, wlogCDR_SimpleRTms_v1_rev +
                  wlogCDR_ChoiceRTms_v1_rev +
                  wlogCDR_DigitVigRTms_v1_rev +
                  wlogJenson8ChoiceDecisionTimems_v1_rev +
                  wlogJenson4ChoiceDecisionTimems_v1_rev +
                  wlogJenson1ChoiceDecisionTimems_v1_rev ~ .imp)

stripplot(cog_vars3_imp, wlogSCB_ConStroopRTms_v1_rev +
            wlogSCB_InconStroopRTms_v1_rev +
            wlogTMT_Ams_v1_rev +
            wlogTMT_Bms_v1_rev + wWASI_vocabrawscore +
            wWASImatrixreasoningrawscore ~ .imp)


stripplot(cog_vars3_imp, wCDR_ImmWRCorrect_v1 +
            wCDR_DelayedWRCorrect_v1 +
            wSerial3sAcc_v1 +
            wSerial7sAcc_v1 + wMMSE + 
            wCDR_NumericWMOverallAcc_v1  ~ .imp)


```


### CFA model 

Here are the data used to fit the CFA model: 

```{r}
varTable(imp_dat)[,c(1,3,4,7:9)]
```

The CFA model is fit using robust maximum likelihood: 

```{r}
options(width = 400)

# define CFA model 

model <- 
  
' # model 

Gt =~  wlogCDR_SimpleRTms_v1_rev +
       wlogCDR_ChoiceRTms_v1_rev +
       wlogCDR_DigitVigRTms_v1_rev +
       wlogJenson8ChoiceDecisionTimems_v1_rev +
       wlogJenson4ChoiceDecisionTimems_v1_rev +
       wlogJenson1ChoiceDecisionTimems_v1_rev
       
Gs =~  wlogSCB_ConStroopRTms_v1_rev +
       wlogSCB_InconStroopRTms_v1_rev +
       wlogTMT_Ams_v1_rev +
       # cross loading
       wlogTMT_Bms_v1_rev +
       wlogCDR_ChoiceRTms_v1_rev 

       
Glr_m6 =~ wCDR_ImmWRCorrect_v1 +
       wCDR_DelayedWRCorrect_v1 
       

Gwm_ac =~ wSerial3sAcc_v1 +
        wSerial7sAcc_v1 +
        wCDR_NumericWMOverallAcc_v1 +
        wlogTMT_Bms_v1_rev

Gf =~ wWASImatrixreasoningrawscore

Gc =~ wMMSE +
      wWASI_vocabrawscore


# correlated residuals
wlogTMT_Ams_v1_rev ~~ wlogTMT_Bms_v1_rev
wSerial3sAcc_v1 ~~ wSerial7sAcc_v1
wlogJenson4ChoiceDecisionTimems_v1_rev ~~ wlogJenson1ChoiceDecisionTimems_v1_rev
wlogJenson8ChoiceDecisionTimems_v1_rev ~~ wlogJenson1ChoiceDecisionTimems_v1_rev
wlogJenson8ChoiceDecisionTimems_v1_rev ~~ wlogJenson4ChoiceDecisionTimems_v1_rev
'

# fit model

test1 <- cfa(model, data = imp_dat, estimator = "MLR", std.lv = F)


# results

summary(test1, fit.measures=T, standardized = T)

```

Extract factor scores from the fitted model for use in later regression models:
 

```{r}

preds <- as_tibble(lavPredict(test1, newdata = imp_dat, type = "lv",
                    method = "Bartlett")) # predict factor scores 

new <- cbind(imp_dat, preds) 

# save only factor scores and IDvar to merge back into data

new <- new %>% select(IDno, Gt:Gc)

arcli2 <- arcli %>% full_join(new, by = "IDno")

write_sav(arcli2, "arcli with CHC.sav") # write output to SPSS file 
```